Nengo: A Python tool for building large-scale functional brain models

### 2013-10-02 nengo.tex
Edits from Bryan's feedback

An \texttt{Ensemble} is
a population of neurons
that represents information
in the form of a real-valued vector.
An ensemble must be given as arguments
its dimensionality
(i.e., the length of the vector it represents),
and an object that describes
the population of neurons.
For example, \texttt{nengo.LIF(50, tau\_ref=0.002)}
describes a population
of 50 leaky integrate-and-fire neurons
with a refractory time of 0.002 seconds.
Neurons are defined symbolically
so that each simulator can compute
this population's nonlinear function
however it sees fit.
The neuron model used by the population
is easily changed, as long as the simulator
is aware of the neuron model description object.

Other attributes of the \texttt{Ensemble},
such as its encoding weights,
the maximum firing rate of its neurons,
and so on, can also be specified
either as keyword arguments
to the \texttt{Ensemble} constructor,
or by setting an attribute on the instantiated object.
While an ensemble makes a hypothesis
about the information being represented by neurons,
these additional attributes
also allow modellers to ensure that
the ensemble stays
within neurobiological constraints.
If these attributes are not set,
then they will be randomly selected
from distributions that describe
neocortical pyramidal cells.

@@

A \texttt{Node} tracks any information
that is not represented by an ensemble.
In the most common case,
nodes provide input signals
that drive ensembles of sensory neurons.
In a more general sense,
nodes represent the experimental environment
in which a neural model exists.

A node computes an arbitrary function
of its inputs directly.
Possible input signals include
the simulator timestep,
the decoded output of an ensemble,
or the output of another node.
However, unlike ensembles,
there are no constraints on the type
of function that the node computes.
A node can track any number of variables internally,
and use the state of those variables
when computing its function.
It can interact directly with hardware,
interface with other programs
using shared memory or sockets,
and so on.

Nodes allow many of the special-purpose
routines that are necessary in real-world
models to be integrated
as core components of a Nengo model.
This makes Nengo models more explicit,
and enables simulators
to control and optimize node execution.

@@

The most important type of connection
in Nengo is the \texttt{DecodedConnection}.
This connection implements
the NEF's transformation principle.
In other words, the \texttt{DecodedConnection}
allows ensembles to project
the information they're representing---or
a transformation of that information---to
another ensemble or node.
This functionality is what enables Nengo models
to appear conceptual,
even though the underlying implementation
can translate that connection
to connection weights.

@@

A probe marks that a signal
will be recorded
throughout the simulation.
Nengo objects and connections
each have several signals that can be probed.
An ensemble's decoded value
and its neurons' spikes
can be probed, among others.

@@

Many models in theoretical neuroscience
are based on attractor networks.
The NEF has been used in the past
to implement many different types of
attractor networks \cite{TODO}.
Figure~\ref{fig:lorez} depicts
a Nengo implementation of the Lorenz chaotic attractor
with a single ensemble
composed of 2000 leaky integrate-and-fire neurons.

>>>>

An \texttt{Ensemble} is
a population of neurons
that represents information
in the form of a real-valued vector.
An ensemble must be given as arguments
its dimensionality
(i.e., the length of the vector it represents),
and an object that describes
the population of neurons.
For example, \texttt{nengo.LIF(50, tau\_ref=0.002)}
describes a population
of 50 leaky integrate-and-fire neurons \cite{TODO} % LIF cite
with a refractory time of 0.002 seconds.
The \texttt{nengo.LIF} class defines
the parameters of the LIF neurons symbolically
so that each simulator can compute
the LIF nonlinearity efficiently.
The neuron model used by the ensemble
is changed by passing in a different symbolic neuron object;
however, the simulator used must be aware
of that neuron model.

Other attributes of the \texttt{Ensemble},
such as its encoding weights,
the maximum firing rate of its neurons,
and so on, can also be specified
either as keyword arguments
to the \texttt{Ensemble} constructor,
or by setting an attribute on the instantiated object.
While an ensemble makes a hypothesis
about the information being represented by neurons,
these additional attributes
allow modellers to fit
neural parameters to \textit{in vivo}
electrophysiology data.
If these attributes are not set,
Nengo attempts to maintain
neurobiological constraints
by selecting neural parameters
from distributions consistent with
neocortical pyramidal cells.

@@

A \texttt{Node} contains a user-defined
Python function that directly calculates
the node's outputs from its inputs at each timestep.
Available inputs include
the simulator timestep,
the decoded output of an ensemble,
or the output of another node.
However, unlike ensembles,
there are no constraints on the type
of function that the node computes.
A node can track any number of variables internally,
and use the state of those variables
when computing its function.
It can interact directly with hardware,
interface with other programs
using shared memory or sockets,
and so on.

A node tracks any information
that is not represented by an ensemble.
As a simple example,
a node can be used to model sensory stimuli
that are predefined functions of time.
As a more sophisticated example,
a node can be used to model
a complex experimental environment that
both provides input to the neural model
and responds to the neural model's output.
Nodes allow many of the special-purpose
routines that are necessary in real-world
models to be integrated
as core components of a Nengo model.
This makes Nengo models more explicit,
and enables simulators
to control and optimize node execution.

@@

The most important type of connection
in Nengo is the \texttt{DecodedConnection}.
This connection implements
the NEF's transformation principle.
In other words, the \texttt{DecodedConnection}
allows ensembles to project
the information they're representing---or
a transformation of that information---to
another ensemble or node.
This functionality is what enables Nengo models
to appear conceptual,
even though the underlying implementation
can translate that connection
to synaptic weights.

@@

A \texttt{Probe} marks that a signal
will be recorded
throughout the simulation.
Nengo models contain many variables
that change over time,
including membrane potentials,
spike events, and represented variables.
It is resource intensive to store the values of
large numbers of variables
at each time step, and it is also not necessary,
as typically only a small fraction
of these variables are analyzed after a simulation.
The modeller chooses which variables to
record by connecting a probe to an object.

@@

Many models in theoretical neuroscience
are based on attractor networks \cite{TODO}. % Amit, Rolls, Deco
The NEF has been used in the past
to implement many different types of
attractor networks \cite{TODO}.
Figure~\ref{fig:lorez} depicts
a Nengo implementation of the Lorenz chaotic attractor
with a single ensemble
composed of 2000 leaky integrate-and-fire neurons.
